Multi-level Selective Deduplication for VM Snapshots in Cloud Storage

### 2012-05-06 vm_snapshot_sim/2012arch.tex
fixed some  typos

%\begin{figure}[htbp]
%  \centering
%  %\epsfig{file=images/dedup_process.eps, height=2in, width=2.66in}
%  \epsfig{file=images/dedup_process.eps,  width=3.2in}
%  \caption{An example  of multi-level deduplication process.}
%  \label{fig:arch}
%\end{figure}
\begin{figure}[htbp]
  \centering
  %\epsfig{file=images/dedup_process.eps, height=2in, width=2.66in}
  \epsfig{file=images/snapshot_representation.eps,  width=3.8in}
  \caption{An example of snapshot representation.}
  \label{fig:snapshot}
\end{figure}

@@

At the runtime, the CDS index resides in a distributed  memory lookup table  
implemented using Memcached~\cite{memcached} to leverage the aggregated memory in a cluster.
Th usage of memory at each machine is small and thus  this scheme  does not
lead to  a memory resource contention with the existing cloud services.
%For example, in our current implementation setting at Aliyun, we use up to  150MB memory per node during night backup
%including CDS and other metadata for all levels of deduplication. Most of memory resource (48GB per node 
%in our experiment) is used for existing  applications.
CDS raw data stored  in the distributed file system
has multiple copies in different machines for the purpose of fault tolerance and 
while providing high read throughput.  
%The data of CDS is stored in a special block store in the distribued file system.

@@

The CDS collected from user-related data is generally proportional to the data size.
As discussed in Section~\ref{sect:exper}, selecting about 1\% of data (after level 1 and level 2 deduplication)
can cover about 38\% of data blocks. Consider we allow maximum 25 VMs per machine, each VM has
about 30GB of user-related data, having 10 snapshots in its snapshot store and the data change ratio during
each backup is 10\%, this translates to 15GB CDS data per machine. 
Consider each CDS hash index entry cost about 40 bytes and the average block size is 4KB, this leads to
a 1:100 ratio so the 
memory cost by CDS hash index is about 150MB per machine.
On the other hand, the CDS from OS-related data is only relvant to the number of OS releases.
Our experiment on 7 major OS releases shows that about 100GB of data is never modified, and we expect
it won't grow over 200GB in the near future. So it would cost the entire cluster 2GB of memory to
store its hash index, or 20MB per machine on a 100 node cluster. On average each OS disk has about 10GB
of data that can be completely eliminated in this way.
\comments{
As the number of VMs increases, the popularity of common blocks for data disks changes.
But this increase has a small impact on OS block popularity because 
there are a limited number of OS releases used by VMs. We let CDS cover the common OS blocks first and
our CDS selection strategy is designed as follows.

>>>>

%\begin{figure}[htbp]
%  \centering
%  %\epsfig{file=images/dedup_process.eps, height=2in, width=2.66in}
%  \epsfig{file=images/dedup_process.eps,  width=3.2in}
%  \caption{An example  of multi-level deduplication process.}
%  \label{fig:arch}
%\end{figure}
\begin{figure}[htbp]
  \centering
  %\epsfig{file=images/dedup_process.eps, height=2in, width=2.66in}
  \epsfig{file=images/snapshot_representation.eps,  width=3.7in}
  \caption{An example of snapshot representation.}
  \label{fig:snapshot}
\end{figure}

@@

At the runtime, the CDS index resides in a distributed  memory lookup table  
implemented using Memcached~\cite{memcached} to leverage the aggregated memory in a cluster.
The usage of memory at each machine is small and thus  this scheme  does not
lead to  a memory resource contention with the existing cloud services.
%For example, in our current implementation setting at Aliyun, we use up to  150MB memory per node during night backup
%including CDS and other metadata for all levels of deduplication. Most of memory resource (48GB per node 
%in our experiment) is used for existing  applications.
CDS raw data stored  in the distributed file system
has multiple copies in different machines for the purpose of fault tolerance and 
while providing high read throughput.  
%The data of CDS is stored in a special block store in the distribued file system.

@@

The CDS collected from user-related data is generally proportional to the data size.
As discussed in Section~\ref{sect:exper}, selecting about 1\% of data (after level 1 and level 2 deduplication)
can cover about 38\% of data blocks. Consider we allow maximum 25 VMs per machine, each VM has
about 30GB of user-related data, having 10 snapshots in its snapshot store and the data change ratio during
each backup is 10\%, this translates to 15GB CDS data per machine. 
Consider each CDS hash index entry cost about 40 bytes and the average block size is 4KB, this leads to
a 1:100 ratio so the 
memory cost by CDS hash index is about 150MB per machine.
On the other hand, the CDS from OS-related data is only relevant to the number of OS releases.
Our experiment on 7 major OS releases shows that about 100GB of data is never modified, and we expect
it won't grow over 200GB in the near future. So it would cost the entire cluster 2GB of memory to
store its hash index, or 20MB per machine on a 100 node cluster. On average each OS disk has about 10GB
of data that can be completely eliminated in this way.
\comments{
As the number of VMs increases, the popularity of common blocks for data disks changes.
But this increase has a small impact on OS block popularity because 
there are a limited number of OS releases used by VMs. We let CDS cover the common OS blocks first and
our CDS selection strategy is designed as follows.

### 2012-05-06 vm_snapshot_sim/2012concl.tex
update
fixed some  typos

\section{Conclusion}
\label{sect:final}
In this paper we propose a multi-level selective deduplication scheme for 
snapshot service in VM cloud. 
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set
effectively  covers a large amount of duplicated data.
Our solution accomplishes the majority of
potential global deduplication saving while still meets stringent cloud resources requirement. 
Evaluation using real user's VM data shows
our solution can accomplish 75\% of what complete global
deduplication can do. 
Compare to today's widely-used snapshot technique, our scheme reduces almost
two-third of snapshot storage cost.
Our scheme uses a very small amount of memory on each node, and leaves
room for additional optimization we are further studying.
%Noted that 6\% is still significant, which is about 24GB per each VM and for a 1000 node Aliyun cluster,
%this is about 600 terabytes.

>>>>

\section{Conclusion}
\label{sect:final}
In this paper we propose a multi-level selective deduplication scheme for 
snapshot service in VM cloud. 
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set
effectively  covers a large amount of duplicated data.
Our solution accomplishes the majority of
potential global deduplication saving while still meets stringent cloud resources requirement. 
Evaluation using real user's VM data shows
our solution can accomplish 75\% of what complete global
deduplication can do. 
Compare to today's widely-used snapshot technique, our scheme reduces almost
two-third of snapshot storage cost.
Finally, our scheme uses a very small amount of memory on each node, and leaves
room for additional optimization we are further studying.
\section*{Acknowledgments}
{\small
We would like to thank Weicai Chen and Shikun Tian from Alibaba for their kind support, 
and the anonymous referees for their comments.
This work was completed while Wei Zhang was a intern at Alibaba.
This work was supported in part by NSF IIS-1118106.
Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and
do not necessarily reflect the views of the National Science Foundation.
}
%Noted that 6\% is still significant, which is about 24GB per each VM and for a 1000 node Aliyun cluster,
%this is about 600 terabytes.

### 2012-05-06 vm_snapshot_sim/2012eval.tex
fixed some  typos

%To see the impact of multi-level deduplication on different OS releases,
%Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
Figure~\ref{fig:oscds} shows the impact of different levels of deduplication for different OS releases.
In this experiment, we tag each block in 350 OS disk snapshots from dataset1 as  ``new''
if this block cannot be deduplicated by our scheme and thus has to be written to the snapshot store;
``CDS''  if this block can be found  in CDS;
``Parent segment'' if  this block is marked unchanged in parent's segment recipe.
``Parent block'' if  this block is marked unchanged in parent's block recipe.
With this tagging, we compute the percentage of deduplication accomplished by each level.
As we can see from Figure~\ref{fig:oscds}, level-1 deduplication accomplishes a large percentage of elimination,
this is because the time interval between two snapshots in our dataset
is quite short and the Aliyun cloud service makes a snapshot  everyday  for each VM.
On the other hand,  CDS still finds lots of duplicates that inner VM deduplication can't find,
contributing about about 10\% of reduction on average.

It is noticable that level-1 deduplication doesn't work well for CentOS, a significant percentage of data is not
eliminated until they reach level-3. It shows that even user upgrade his VM system heavily and frequently
such that data locality is totally lost, those OS-related data can still be identified at level-3.

In gengral we see a stable data reduction ratio for all OS varieties, ranging from 92\% to 97\%, that means
the storage cost of 10 full snapshots combined is still smaller than the original disk size. And compare to 
today's widely used copy-on-write snapshot technique, which is similar to our level-1 deduplication, our
solution cut the snapshot storage cost by 64\%.

@@

One of our biggest advantage is small memory footprint for deduplication, because we only
eliminate a small amount of highly popular data.
we use dataset2 to demostrate the coverage of popular data blocks on user-related data, 
as shown in Figure \ref{fig:userdatacoverage}.
%content blocks used in image files using 90\% of
%the collected dataset.
%Figure \ref{fig:userdatacoverage} shows the cumulative coverage  ratio of  popular common user data blocks. 
The $X$ axis is the rank of cmomon user data blocks, and $Y$ shows how much raw data can be covered given
the size of CDS.
Let $S_i$ and $F_I$ be the size and duplicate count 
of the $i$-th block ranked by its duplicate rank.  Then $Y$ axis is the coverage of
the common dataset covering data items from rank 1 to rank $i$. Namely  
\[
\frac{ \sum_{i=1}^{i} S_i * F_i} {\mbox{Total data size}}.
\]

Thus with about 1\% of blocks on data disks, CDS can cover about 38\% of total data blocks
appeared in in all data snapshots. The corresponding CDS index uses no more than 150MB memory per machine,
which can be easily co-located with other cloud services.

>>>>

%To see the impact of multi-level deduplication on different OS releases,
%Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
Figure~\ref{fig:oscds} shows the impact of different levels of deduplication for different OS releases.
In this experiment, we tag each block in 350 OS disk snapshots from dataset1 as  ``new''
if this block cannot be deduplicated by our scheme and thus has to be written to the snapshot store;
``CDS''  if this block can be found  in CDS;
``Parent segment'' if  this block is marked unchanged in parent's segment recipe.
``Parent block'' if  this block is marked unchanged in parent's block recipe.
With this tagging, we compute the percentage of deduplication accomplished by each level.
As we can see from Figure~\ref{fig:oscds}, level-1 deduplication accomplishes a large percentage of elimination,
this is because the time interval between two snapshots in our dataset
is quite short and the Aliyun cloud service makes a snapshot  everyday  for each VM.
On the other hand,  CDS still finds lots of duplicates that inner VM deduplication can't find,
contributing about 10\% of reduction on average.

It is noticeable that level-1 deduplication doesn't work well for CentOS, a significant percentage of data is not
eliminated until they reach level-3. It shows that even user upgrade his VM system heavily and frequently
such that data locality is totally lost, those OS-related data can still be identified at level-3.

In general we see a stable data reduction ratio for all OS varieties, ranging from 92\% to 97\%, that means
the storage cost of 10 full snapshots combined is still smaller than the original disk size. And compare to 
today's widely used copy-on-write snapshot technique, which is similar to our level-1 deduplication, our
solution cut the snapshot storage cost by 64\%.

@@

One of our biggest advantage is small memory footprint for deduplication, because we only
eliminate a small amount of highly popular data.
we use dataset2 to demonstrate the coverage of popular data blocks on user-related data, 
as shown in Figure \ref{fig:userdatacoverage}.
%content blocks used in image files using 90\% of
%the collected dataset.
%Figure \ref{fig:userdatacoverage} shows the cumulative coverage  ratio of  popular common user data blocks. 
The $X$ axis is the rank of common user data blocks, and $Y$ shows how much raw data can be covered given
the size of CDS.
Let $S_i$ and $F_I$ be the size and duplicate count 
of the $i$-th block ranked by its duplicate rank.  Then $Y$ axis is the coverage of
the common dataset covering data items from rank 1 to rank $i$. Namely  
\[
\frac{ \sum_{i=1}^{i} S_i * F_i} {\mbox{Total data size}}.
\]

Thus with about 1\% of blocks on data disks, CDS can cover about 38\% of total data blocks
appeared in all data snapshots. The corresponding CDS index uses no more than 150MB memory per machine,
which can be easily co-located with other cloud services.

### 2012-05-06 vm_snapshot_sim/2012intro.tex
fixed some  typos

Unlike the previous work dealing with general file-level backup and deduplication, our problem is focused on 
virtual disk image backup. Although we treat each virtual disk  as a file logically, its size is very large.
On the other hand, we need to support parallel backup of a large number of virtual disks in a cloud everyday. 
One key requirement we face at Alibaba Aliyun is that VM snapshot backup should only use a minimal amount of system
resources so that most of resources is kept for regular cloud system services or applications.
Thus our objective is to exploit the characteristics of VM snapshot data and
pursue a cost-effective deduplication solution. 
Another goal  is to decentralize VM snapshot backup and  localize  deduplication as much as possible,
which brings the benefits for increased parallelism  and fault isolation.

@@

With these in mind, we  have developed a decentralized multi-level solution to conduct 
segment-level  and block-level inner-VM  deduplication to localize the deduplication effort when possible.
It then makes cross-VM deduplication by excluding a small number of
popular common data blocks from being backed up. Our study shows that common data blocks
occupy significant amount of storage space while they only take
a small amount of resources to deduplicate.
Seperating deduplication into multi levels effectively accomplish the major space saving goal
compare the global complete deduplication scheme, at the same time it makes
the backup of different VMs to be indepdent for better fault tolerance.

>>>>

Unlike the previous work dealing with general file-level backup and deduplication, our problem is focused on 
virtual disk image backup. Although we treat each virtual disk  as a file logically, its size is very large.
On the other hand, we need to support parallel backup of a large number of virtual disks in a cloud every day. 
One key requirement we face at Alibaba Aliyun is that VM snapshot backup should only use a minimal amount of system
resources so that most of resources is kept for regular cloud system services or applications.
Thus our objective is to exploit the characteristics of VM snapshot data and
pursue a cost-effective deduplication solution. 
Another goal  is to decentralize VM snapshot backup and  localize  deduplication as much as possible,
which brings the benefits for increased parallelism  and fault isolation.

@@

With these in mind, we  have developed a decentralized multi-level solution to conduct 
segment-level  and block-level inner-VM  deduplication to localize the deduplication effort when possible.
It then makes cross-VM deduplication by excluding a small number of
popular common data blocks from being backed up. Our study shows that common data blocks
occupy significant amount of storage space while they only take
a small amount of resources to deduplicate.
Separating deduplication into multi levels effectively accomplish the major space saving goal
compare the global complete deduplication scheme, at the same time it makes
the backup of different VMs to be independent for better fault tolerance.

### 2012-05-06 vm_snapshot_sim/main.tex
fixed some  typos

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{Multi-level Selective Deduplication for VM Snapshots in Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}
\author{
\IEEEauthorblockN{Wei Zhang\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Hong Tang\IEEEauthorrefmark{2}, Hao Jiang\IEEEauthorrefmark{2}, Tao Yang\IEEEauthorrefmark{1}, Xiaogang Li\IEEEauthorrefmark{2}, Rachael Zeng\IEEEauthorrefmark{2}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Dept. of Computer Science, UC Santa barbara. Email: \{wei, tyang\}@cs.ucsb.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Alibaba Inc. Email: \{hongtang, haojiang, xiaogang, rachael\}@alibaba-inc.com}
%Wei Zhang$^{\star\dagger}$, Hong Tang$^\dagger$, Hao Jiang$^\dagger$, Tao Yang$^\star$, 
%Xiaogang Li$^\dagger$, and Rachael Zeng$^\dagger$ \\
%   {\normalsize $^\star$UC Santa Barbara}\\
%   {\normalsize$^\dagger$Alibaba.com}
}

>>>>

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{Multi-level Selective Deduplication for VM Snapshots in Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}
\author{
\IEEEauthorblockN{Wei Zhang\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Hong Tang\IEEEauthorrefmark{2}, Hao Jiang\IEEEauthorrefmark{2}, Tao Yang\IEEEauthorrefmark{1}, Xiaogang Li\IEEEauthorrefmark{2}, Yue Zeng\IEEEauthorrefmark{2}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Dept. of Computer Science, UC Santa Barbara. Email: \{wei, tyang\}@cs.ucsb.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Alibaba Inc. Email: \{hongtang, haojiang, xiaogang, rachael\}@alibaba-inc.com}
%Wei Zhang$^{\star\dagger}$, Hong Tang$^\dagger$, Hao Jiang$^\dagger$, Tao Yang$^\star$, 
%Xiaogang Li$^\dagger$, and Rachael Zeng$^\dagger$ \\
%   {\normalsize $^\star$UC Santa Barbara}\\
%   {\normalsize$^\dagger$Alibaba.com}
}
